---
title: "Categorical EDA Quiz Worksheet"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

## Setup

### Packages Used

```{r}
library(tidyverse)
library(resampledata)
```

### Datasets Used

#### NFL Players Data

The `NFLplayers2016.csv` dataset contains data about all National Football League (NFL) players in 2016. You can read more about the variables on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html). 

```{r}
nfl <-
  read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv") %>%
  mutate(Playtime = factor(Playtime,
    levels = c(
      "Rookie",
      "1-4 Yrs",
      "5-8 Yrs",
      "9-12 Yrs",
      "13+ Yrs"
    )
  ))
glimpse(nfl)
```

#### Smallpox Data

[William Watson](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4109147/) (1717-1787) conducted one of the earliest known examples of the type of research we now call a clinical trial. The `smallpox.csv` dataset contains data from three studies in which he tested various pre-treatments that were believed to improve the effectiveness of smallpox inoculations. 

```{r}
pox <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/smallpox.csv")
glimpse(pox)
```

## Problems

### NFL Players 2016

#### Table of `Side`

Create a table of frequencies and relative frequencies for the variable `Side`.

```{r}

```

#### Table of `PositionGrp`

Create a table of frequencies and relative frequencies for the variable `PositionGrp`. Use `arrange()` to sort the rows in descending order (i.e., the most frequent category at the top).

```{r}

```

#### Table of `Playtime`

Tabulate frequencies, cumulative frequencies, relative frequencies, and cumulative relative frequencies for the variable `Playtime`. 

```{r}

```

#### Tables of `Side` x `Playtime`

```{r}

```

##### Frequencies

Create a contingency table of counts for `Side` by `Playtime`.

```{r}

```

##### Row Relative Frequencies

Create a contingency table of the relative frequencies of `Playtime` within each category (row) of `Side`.

```{r}

```

#### Top Five Player `College`s

Find the top five colleges among NFL players in 2016.

```{r}

```

#### Questions

About _____ % of NFL players were classified as special teams. 

Answer:

The number of players on offense was _____ the number of players on defense.

(A) substantially more than
(B) moderately more than
(C) approximately the same as
(D) moderately less than
(E) substantially less than

Players in the _____ group(s) seem to be more likely than other players to have a long NFL career.

(A) defense
(B) offense
(C) special teams
(D) defense and offense
(E) defense and special teams
(F) offense and special teams

### Smallpox Data

#### Table of `experiment` x `pretreatment`

Create a contingency table of counts for `experiment` by `pretreatment`.

```{r}

```

#### Questions

Which pretreatments were used in Study 1? 

(A) Mercury Only
(B) Mercury and Jalap
(C) Senna and Rose Syrup
(D) Mercury Only *and* Mercury and Jalap
(E) Mercury Only *and* Senna and Rose Syrup
(F) Mercury and Jalap *and* Senna and Rose Syrup
(G) Mercury Only, Mercury and Jalap, *and* Senna and Rose Syrup
(H) none, it was a control study only

Which pretreatments were used in Study 2? 

(A) Mercury Only
(B) Mercury and Jalap
(C) Senna and Rose Syrup
(D) Mercury Only *and* Mercury and Jalap
(E) Mercury Only *and* Senna and Rose Syrup
(F) Mercury and Jalap *and* Senna and Rose Syrup
(G) Mercury Only, Mercury and Jalap, *and* Senna and Rose Syrup
(H) none, it was a control study only

Which pretreatments were used in Study 3? 

(A) Mercury Only
(B) Mercury and Jalap
(C) Senna and Rose Syrup
(D) Mercury Only *and* Mercury and Jalap
(E) Mercury Only *and* Senna and Rose Syrup
(F) Mercury and Jalap *and* Senna and Rose Syrup
(G) Mercury Only, Mercury and Jalap, *and* Senna and Rose Syrup
(H) none, it was a control study only
 
<hr>
```{r}
sessionInfo()
```
