---
title: "Numerical EDA Quiz Worksheet"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_depth: 5
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
<hr>

## Setup

### Packages Used

```{r}
library(tidyverse)
```

### Datasets Used

#### NFL Players Data

The `NFLplayers2016.csv` dataset contains data about all National Football League (NFL) players in 2016. You can read more about the variables on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/NFLplayers2016.html). 

```{r}
nfl <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/NFLplayers2016.csv")
glimpse(nfl)
```

#### Monopoly Data

The `monopolygame.csv` dataset contains data about properties in the game of Monopoly. You can read more about the variables on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/monopoly.html) 

```{r}
game <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/monopolygame.csv")
glimpse(game)
```

#### Smallpox Data

[William Watson](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4109147/) (1717-1787) conducted one of the earliest known examples of the type of research we now call a clinical trial. The `smallpox.csv` dataset contains data from three studies in which he tested various pre-treatments that were believed to improve the effectiveness of smallpox inoculations. 

```{r}
pox <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/smallpox.csv")
glimpse(pox)
```

## Problems

### NFL Players Data

#### Descriptive Summaries

Summarize n, mean, standard deviation, median, and IQR for `Weight` grouped by `Side` to create **Table 1**.

```{r}
nfl %>% 
  group_by(Side) %>% 
  summarize(n     = n(),
            xbar  = mean(Weight),
            s     = sd(Weight),
            xmed  = median(Weight),                  
            iqr   = IQR(Weight))
```

Summarize the deciles for `Weight` (10th, 20th, 30th, ..., 90th quantiles) grouped by `Side` to create **Table 2**.

```{r}
nfl %>% 
  group_by(Side) %>% 
  summarize(
    Quantile = c("10th", "20th", "30th", "40th", "50th", "60th", "70th", "80th", "90th"),
    Deciles = quantile(Weight, probs = seq(0.1, 0.9, 0.1))) %>% 
  pivot_wider(names_from = Side, values_from = Deciles)
```

Compute the lower and upper fences for `Weight` grouped by `Side` to create **Table 3**.

```{r}
nfl %>% 
  group_by(Side) %>%
  summarize(LF = quantile(Weight, 0.25) - 1.5 * IQR(Weight),
            UF = quantile(Weight, 0.75) + 1.5 * IQR(Weight))
```

Create density plots of `Weight` for each category of `Side` to observe the distribution shape.

```{r}
nfl %>% 
  filter(Side == "Offense") %>% 
  ggplot(aes(x = Weight)) +
  geom_density()

nfl %>% 
  filter(Side == "Defense") %>% 
  ggplot(aes(x = Weight)) +
  geom_density()

nfl %>% 
  filter(Side == "Special Teams") %>% 
  ggplot(aes(x = Weight)) +
  geom_density()
```

```{r}
nfl %>% 
  ggplot(aes(x = Weight)) +
  geom_density() +
  facet_grid(. ~ Side)
```

#### Questions

A player weighing 225 pounds would be heavier than about _____  % of offense and defense players, whereas they'd be heavier than about _____ % of special teams players.

Answer 1:

Answer 2:

Based on the this initial exploration, the _____ group(s) is/are most likely to need a more detailed breakdown and in-depth investigation to really understand the distribution of weight among those players. 

(A) defense
(B) offense
(C) special teams
(D) defense and offense
(E) defense and special teams
(F) offense and special teams

### Monopoly Data

#### Descriptive Summaries

Summarize n, mean, and standard deviation for property `cost` grouped by `type` to create part of **Table 4**.

```{r}

```

Summarize n, mean, and standard deviation for property `cost` (all the properties) to create part of **Table 4**.

```{r}

```

Create plots to visualize the shapes of the distributions.

```{r}
ggplot(game, aes(x = cost)) +
  geom_density()

ggplot(game, aes(x = cost)) +
  geom_histogram(bins = 8)

game %>% 
  filter(type == "street") %>% 
  ggplot(aes(x = cost)) +
  geom_density()
```

#### Questions

The group with the greatest variability in prices is _____ properties.

(A) street
(B) railroad
(C) utility
(D) all

The distribution of street property costs is _____ .

(A) strongly left skewed
(B) somewhat left skewed
(C) approximately symmetric
(D) somewhat right skewed
(E) strongly right skewed
(F) not applicable because there is no variability

The distribution of railroad property costs is _____ .

(A) strongly left skewed
(B) somewhat left skewed
(C) approximately symmetric
(D) somewhat right skewed
(E) strongly right skewed
(F) not applicable because there is no variability
 
The distribution of utility property costs is _____ .

(A) strongly left skewed
(B) somewhat left skewed
(C) approximately symmetric
(D) somewhat right skewed
(E) strongly right skewed
(F) not applicable because there is no variability

The distribution of all property costs is _____ .

(A) strongly left skewed
(B) somewhat left skewed
(C) approximately symmetric
(D) somewhat right skewed
(E) strongly right skewed
(F) not applicable because there is no variability

### Smallpox Data

#### Descriptive Summaries

For Experiment 1 only, compute the five-number summary for `pocks` by `pretreatment` group to create **Table 5**.

```{r}
pox %>% 
  filter(experiment == 1) %>% 
  group_by(pretreatment) %>% 
  summarize(n     = n(),
            min   = min(pocks),
            Q1    = quantile(pocks, 0.25),
            xmed  = median(pocks),                  
            Q3    = quantile(pocks, 0.75),
            max   = max(pocks))
```

Calculate the upper fence for all observations that received no pretreatment, regardless of experiment number. 

```{r}
pox %>% 
  filter(pretreatment == "none") %>%
  summarize(UF = round(quantile(pocks, 0.75) + 1.5 * IQR(pocks), 2))
```

#### Questions

Based on this criteria (i.e., the upper fence), the max number of pocks observed in the control group in Experiment 1 _____ be an outlier.

(A) would
(B) would not

<hr>
```{r}
sessionInfo()
```
