---
title: "Correlation & Regression"
author: "Dr. Thomley"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Packages Used

```{r packages}
library(tidyverse)
library(moderndive)
```

## DataSets Used

### Anscombe Data

```{r}
glimpse(anscombe)
```

### Anthropometric Data

You can read about the dataset on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/anthropometric.html).

```{r datasets}
anthro <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/anthropometric.csv")
glimpse(anthro)
```

```{r}
set.seed(1)
anthro_sample <- sample_n(anthro, 20)
glimpse(anthro_sample)
```

### Monopoly Game Data

You can read about the dataset on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/monopoly.html)

```{r}
monopoly <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/monopolygame.csv")
glimpse(monopoly)
```

## Examples

### Scatterplot of Ideal Height vs. Height

```{r}
ggplot(anthro, aes(x = height, y = ideal)) +
  geom_point()
```

## Correlation for Ideal vs. Height

```{r}
anthro %>% 
  summarize(r = cor(height, ideal, use = "pairwise.complete.obs"))
```

Is this strong? How large is "strong"? Depends on the discipline.

For this class, we'll divide the values into thirds.

* 0 to about .33 = low or weak (same for negative corrs)
* about 0.33 to about 0.67 = moderate
* about 0.67 to 1 = high or strong

### Least Squares Line --- Visualization

```{r}
ggplot(anthro, aes(x = height, y = ideal)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) # lm is linear model
```

```{r}
lm(ideal ~ height, data = anthro) # y ~ x
```

```{r}
square_model <- lm(armspan ~ height, data = anthro_sample)
summary(square_model)
```


### Scatterplot of Cost by Spaces from Go for Monopoly

```{r}
ggplot(monopoly, aes(x = spaces, y = cost, color = type, shape = type)) +
  geom_point(size = 4) +
  labs(title = "Purchase Prices of Properties in the Game of Monopoly",
       x = "Spaces from Go", 
       y = "Purchase Price",
       color = "Property Type",    # sets the legend title for the color
       shape = "Property Type") +  # sets the legend title for the shape
  scale_color_brewer(palette = "Dark2")
```

```{r}
ggplot(monopoly, aes(x = spaces, y = cost, color = type, shape = type)) +
  geom_point(size = 4) +
  labs(title = "Purchase Prices of Properties in the Game of Monopoly",
       x = "Spaces from Go", 
       y = "Purchase Price",
       color = "Property Type",    # sets the legend title for the color
       shape = "Property Type") +  # sets the legend title for the shape
  scale_color_brewer(palette = "Dark2") +
  geom_smooth(method = "lm", se = FALSE)
```

### Correlation for All Properties

```{r}
monopoly %>% 
  summarize(r = cor(spaces, cost))
```

### Correlation by Property Type

```{r}
monopoly %>% 
  group_by(type) %>% 
  summarize(r = cor(spaces, cost))
```

The correlation for streets alone is larger than for all properties together. Why? Railroads and utilities are outliers with regard to the pattern established by spaces and streets --- spaces has no relationship with railroad and utility proces.

```{r}
lm(cost ~ spaces, data = monopoly)
```



### Anscombe Data

Covariance is approximately 5.5 for each pair.

```{r}
anscombe %>% 
  summarize(cov1 = cov(x1, y1),
            cov2 = cov(x2, y2),
            cov3 = cov(x3, y3),
            cov4 = cov(x4, y4))
```

Correlation is approximately 0.816 for all pairs.

```{r}
anscombe %>% 
  summarize(r1 = cor(x1, y1),
            r2 = cor(x2, y2),
            r3 = cor(x3, y3),
            r4 = cor(x4, y4))
```

If we relied on only the calculations, we'd say the strength of the *linear* relationshps are the same. BUT...

***
```{r}
sessionInfo()
```