---
title: "EDA ⯈ Categorical Data Worksheet"
author: "YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Packages Used

```{r packages}
library(tidyverse)
library(waffle)
```

## DataSets Used

### Popular Kids Study

You can read more about this dataset on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r datasets}
kids <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/popular.csv")
```

### STT3850 Students

This dataset contains information about students in Dr. Thomley’s past STT3850 classes.

```{r}
students <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv")
```

## Tabulating Data

### Frequency 

```{r}

```

### Relative Frequency

```{r}

```

### Cumulative Frequency

```{r}

```

### Using Base R

## Plotting Data

### Barplot for One Variable

```{r}

```

### Barplot for Two Variables

```{r}

```

### Barplot for Three Variables

```{r}

```

### Waffle Plot

```{r}

```

## Creating Factors

### *create your own subheader here*

```{r}

```

***
```{r}
sessionInfo()
```
