---
title: "EDA ⯈ Numerical Data Worksheet"
author: "YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Packages Used

```{r packages}
library(tidyverse)
library(resampledata) # needed for wells dataset
```

## DataSets Used

### Mileage Data

You can read more about this dataset on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r datasets}
mileage <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
```

### Geyser Data

This dataset contains information about students in Dr. Thomley’s past STT3850 classes.

```{r}
geyser <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv")
```

### Wells Data

```{r}
wells <- Bangladesh
```

### Happy Face Data

```{r}
geyser <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/happyface.csv")
```

## Exploratory Data Analysis (EDA)

### Mileage Data

#### Plots

```{r}

```

#### Summaries

```{r}

```

### Geyser Data

#### Plots

```{r}

```

#### Summaries

```{r}

```

### Wells Data

#### Plots

```{r}

```

#### Summaries

```{r}

```

### Happy Face Data

#### Plots

```{r}

```

#### Summaries

```{r}

```

***
```{r}
sessionInfo()
```
