---
title: "EDA ⯈ Numerical Data Worksheet"
author: "YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Packages Used

```{r packages}
library(tidyverse)
library(resampledata) # needed for wells dataset
```

## DataSets Used

### Mileage Data

```{r datasets}
mileage <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/gasmileage.csv")
glimpse(mileage)
```

### Geyser Data

This dataset contains information about students in Dr. Thomley’s past STT3850 classes.

```{r}
geyser <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/oldfaithful.csv")
glimpse(geyser)
```

### Wells Data

```{r}
wells <- Bangladesh
glimpse(wells)
```

### Happy Face Data

```{r}
happyf <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/happyface.csv")
glimpse(happyf)
```

## Exploratory Data Analysis (EDA)

### Mileage Data

#### Plots

Histogram and Frequency Polygon

```{r}
ggplot(mileage, aes(x = mpg)) +
  geom_histogram(bins = 15, fill = "orange", color = "white") +
  geom_freqpoly(bins = 15, color = "blue")
```

Density Plot

```{r}
ggplot(mileage, aes(x = mpg)) +
  geom_density()
```

#### Summaries

Using Base Code

```{r}
mean(mileage$mpg)
sd(mileage$mpg)
fivenum(mileage$mpg)
quantile(mileage$mpg, probs = seq(from = 0.1, to = 0.9, by = 0.1))
```

Using `summarize()` 

Since the data are symmetric, the mean and median are about the same value. Since the data Are roughly bell-shaped, the range is equal to about 6 standard deviations (mean +/- 3SD).

```{r}
mileage %>%
  summarize(
    n = n(),
    xbar = mean(mpg),
    s = sd(mpg),
    xmin = min(mpg),
    Q1 = quantile(mpg, 0.25),
    med = median(mpg),
    Q3 = quantile(mpg, 0.75),
    xmax = max(mpg),
    iqr = IQR(mpg),
    rng = diff(range(mpg)),
    total = sum(mpg)
  )
```

We can use just the stats we need...

```{r}
mileage %>%
  summarize(
    n = n(),
    xbar = mean(mpg),
    s = sd(mpg)
  )
```

What are the fences for the data?

```{r}
mileage %>%
  summarize(
    LF = quantile(mpg, 0.25) - 1.5 * IQR(mpg),
    UF = quantile(mpg, 0.75) + 1.5 * IQR(mpg)
  )
```

```{r}
mpg_fences <- mileage %>%
  summarize(LF = quantile(mpg, 0.25) - 1.5 * IQR(mpg),
            UF = quantile(mpg, 0.75) + 1.5 * IQR(mpg))

LF <- mpg_fences %>% 
  pull(LF)

UF <- mpg_fences %>% 
  pull(UF)
```

```{r}
ggplot(mileage, aes(x = mpg)) +
  geom_boxplot()
```

What are the values of the outliers? They are the min and max from our descriptive stats.

```{r}
mileage %>% 
  filter(mpg < LF | mpg > UF)
```



### Geyser Data

#### Plots

```{r}
geyser <- geyser %>%
  mutate(group = ifelse(erupt_time < 3.1, "short", "long"))

ggplot(geyser, aes(x = erupt_time, fill = group)) +
  geom_histogram(color = "white")
```

```{r}
ggplot(geyser, aes(x = wait_time)) + 
  geom_boxplot() +
  scale_y_continuous(breaks = NULL) +
  labs(title = "Wait time between Old Faithful eruptions",
       x = "time in minutes",
       y = NULL,
       caption = "source: oldfaithful.csv dataset")
```

```{r}
ggplot(geyser, aes(x = erupt_time, y = group)) + 
  geom_boxplot() +
  labs(title = "Wait time between Old Faithful eruptions",
       x = "time in minutes",
       y = NULL,
       caption = "source: oldfaithful.csv dataset")
```

#### Summaries

```{r}
geyser %>% 
  summarize(n     = n(),
            xbar  = mean(erupt_time),
            s     = sd(erupt_time),
            xmin  = min(erupt_time),
            Q1    = quantile(erupt_time, 0.25),
            med   = median(erupt_time),                  
            Q3    = quantile(erupt_time, 0.75),
            xmax  = max(erupt_time),
            iqr   = IQR(erupt_time),
            rng   = diff(range(erupt_time)),
            total = sum(erupt_time))
```

```{r}
geyser %>% 
  group_by(group) %>% 
  summarize(n     = n(),
            xbar  = mean(erupt_time),
            s     = sd(erupt_time),
            xmin  = min(erupt_time),
            Q1    = quantile(erupt_time, 0.25),
            med   = median(erupt_time),                  
            Q3    = quantile(erupt_time, 0.75),
            xmax  = max(erupt_time),
            iqr   = IQR(erupt_time),
            rng   = diff(range(erupt_time)),
            total = sum(erupt_time))
```


### Wells Data

#### Plots

```{r}
ggplot(wells, aes(x = Arsenic)) +
  geom_histogram(aes(y = ..density..), bins = 25, fill = "orange") +
  geom_density(color = "blue")
```

#### Summaries

```{r}
wells %>% 
  summarize(n     = n(),
            xbar  = mean(Arsenic),
            s     = sd(Arsenic),
            xmin  = min(Arsenic),
            Q1    = quantile(Arsenic, 0.25),
            med   = median(Arsenic),                  
            Q3    = quantile(Arsenic, 0.75),
            xmax  = max(Arsenic),
            iqr   = IQR(Arsenic),
            rng   = diff(range(Arsenic)),
            total = sum(Arsenic))

mean(wells$Arsenic, trim = 0.05)  # 5% trimmed mean
```

```{r}
wells %>% 
  summarize(n     = n(),
            xmed  = median(Chlorine, na.rm = TRUE))

wells %>% 
  summarize(n     = n(),
            iqr   = IQR(Chlorine, na.rm = TRUE))
```

```{r}
ggplot(wells, aes(x = Arsenic)) + 
  geom_boxplot() +
  scale_y_continuous(breaks = NULL) +
  labs(title = "Arsenic levels in 271 sampled wells in Bangladesh",
       x = "parts per billion",
       y = NULL,
       caption = "source: Bangladesh.csv dataset")
```

```{r}
arsenic_fences <- wells %>%
  summarize(LF = quantile(Arsenic, 0.25) - 1.5 * IQR(Arsenic),
            UF = quantile(Arsenic, 0.75) + 1.5 * IQR(Arsenic))
arsenic_fences
```

```{r}
LF <- arsenic_fences %>% 
  pull(LF)

UF <- arsenic_fences %>% 
  pull(UF)

wells %>% 
  filter(Arsenic < LF | Arsenic > UF) %>% 
  arrange(desc(Arsenic))
```

Remove outliers...filter for points that lie between the fences...greater than or = LF *and* less than or = UF.

```{r}
wells %>% 
  filter(Arsenic >= LF & Arsenic <= UF)

wells %>% 
  filter(Arsenic >= LF, Arsenic <= UF)
```


### Happy Face Data

#### Plots

```{r}

```

#### Summaries

```{r}

```

***
```{r}
sessionInfo()
```
