---
title: "EDA ⯈ Categorical Data Worksheet"
author: "Jill E. Thomley"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

## Packages Used

```{r packages}
library(tidyverse)
library(waffle)
```

## DataSets Used

### Popular Kids Study

You can read more about this dataset on its [description page](https://stat-jet-asu.github.io/Datasets/InstructorDescriptions/popular.html).

```{r datasets}
kids <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/popular.csv")
glimpse(kids)
```

### STT3850 Students

This dataset contains information about students in Dr. Thomley’s past STT3850 classes.

```{r}
students <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv")
glimpse(students)
```

## Tabulating Data

### Frequency 

I want to take the kids dataset *and then* count how many kids are in each grade. The pipe can be read as *and then* in this process. The counts produced by the `count()` function go into a column labeled `n`. The output  of this process is another dataset (tibble). This is good, because we can use other dplyr verbs to manipulate this output.

```{r}
kids %>% 
  count(grade)
```

### Relative Frequency

The way we add a new column or variable to a dataset is with the `mutate()` function. I mathematically manipulate the `n` column into the `rel_freq` column using the function `prop.table()`.

```{r}
kids %>%
  count(grade) %>%
  mutate(rel_freq = prop.table(n))
```

### Cumulative Frequency

Using `mutate()` and the cumulative sum function `cumsum()`, I can manipulate the `n` and `rel_freq` columns to add cumulative frequency and cumulative relative frequency columns.

```{r}
kids %>%
  count(grade) %>%
  mutate(
    cum_freq = cumsum(n),
    rel_freq = prop.table(n),
    cum_rel_freq = cumsum(rel_freq)
  )
```

### New Example!

To plot a different variable from the same dataset, just change the variable name!

What are kids' top goals?

```{r}
kids %>%
  count(top_goal) %>%
  mutate(rel_freq = prop.table(n))
```

Where do kids rank grades in terms of contributing to popularity (1 = most important)?

```{r}
kids %>%
  count(grades_rank) %>%
  mutate(
    cum_freq = cumsum(n),
    rel_freq = prop.table(n),
    cum_rel_freq = cumsum(rel_freq)
  )
```

### Using Base R

## Plotting Data

### Barplot for One Variable

These are the same plot *geometry*, except the *aesthetic* (`aes`) variable is on the x axis in one plot and the y axis in the other. The `+` sign in ggplot works much like the `%>%`---it adds another layer to the plot. 

```{r}
ggplot(kids, aes(x = top_goal)) +
  geom_bar()

ggplot(kids, aes(y = top_goal)) +
  geom_bar()
```

### Barplot for Two Variables

```{r}

```

### Barplot for Three Variables

```{r}

```

### Waffle Plot

```{r}

```

## Creating Factors

### *create your own subheader here*

```{r}

```

***
```{r}
sessionInfo()
```
