---
title: "Lab 01: Who Takes STT3850?" 
subtitle: "EDA ⯈ Categorical Data"
author: "YOUR NAME HERE"
date: 'Last rendered on `r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    code_folding: hide
---

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```
***

Collaborators: *add the names of anyone you worked with here*

## Packages Used

```{r packages}
library(tidyverse)
library(waffle)
```

## DataSet Used

### Load the Data

The `stt3850majors` dataset contains information about students in Dr. Thomley’s past and present STT3850 classes.

```{r}
students <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/Instructor/stt3850majors.csv")
glimpse(students)
```

* `College`: college to which the student belonged (based on major)
* `Year`: student's academic level/year as listed in the Banner system
* `BannerMajor`: student's major as listed in the Banner system
* `CodedMajor`: student's major recoded into four categories
* `Semester`: semester in which the student took STT3850
* `AY`: academic year for `Semester`, e.g., 2014-2015 (F14, S15, Sum 15)

All of Dr. Thomley's sections prior to Spring 2020 were taught in person. Spring 2020 began in person and moved to 100% online. All sections after Spring 2020 have been taught 100% online.

### Prepare the Data

* Filter the data to remove the one student whose `Year` is Graduate Certificate Program (hint: use `!=`).
* Mutate the variable `Year` into a factor so the categories Freshman, Sophomore, Junior, and Senior are in order. 
* Mutate the variable `CodedMajor` into a factor so the categories are in the order shown below.
    * Computer Science
    * Mathematical Sciences
    * Other Sciences
    * Other
* Run `glimpse()` to show the new structure.

```{r}
students <- students %>% 
  filter(...) %>% 
  mutate(...) 
```

## Frequency Tables

### Table 1: Coded Major

Create a table showing the frequency and relative frequency of `CodedMajor`. 

```{r}

```

### Table 2: Class Year

Create a table showing the frequency, relative frequency, cumulative frequency, and cumulative relative frequency of `Year`.

```{r}

```

## Contingency Tables

### Tables 3 and 4: Coded Major by Class Year

Create a contingency table of `CodedMajor` (rows) by `Year` (columns). Use `pivot_wider()` to spread it out for easier interpretation (hint: save the output of your `count()` as an object before using `pivot_wider()`, then use it as input for the next table as well.)

```{r}

```

Create a contingency table of `CodedMajor` (rows) by `Year` (columns) and compute relative frequencies of `Year` grouped by `CodedMajor` (i.e., relative frequencies within each row). Use `pivot_wider()` to spread it out for easier interpretation.

```{r}

```

### Tables 5 and 6: Academic Year by Coded Major

Create a contingency table of `AY` (rows) by `CodedMajor` (columns). Use `pivot_wider()` to spread it out for easier interpretation (hint: save the output of your `count()` as an object before using `pivot_wider()`, then use it as input for the next table as well.)

```{r}

```

Create a contingency table of `AY` (rows) by `CodedMajor` (columns) and compute relative frequencies of `CodedMajor` grouped by `AY` (i.e., relative frequencies within each row). Use `pivot_wider()` to spread it out for easier interpretation.

```{r}

```

## Barplots

### Plot 1: Class Year

Create a barplot of `Year`. Include an informative title and axis labels using `labs()`. Other colors/formatting are optional.

```{r}

```

### Plot 2: Coded Major by Class Year

Create a barplot of `CodedMajor` with each bar filled according to `Year`. Use the `geom_bar()` option `position = "dodge"` to unstack the bars. Include an informative title and axis labels using `labs()`. Other colors/formatting are optional.

```{r}

```

### Plot 3: Academic Year by Coded Major

Create a horizontal barplot of `AY` with each bar filled according to `CodedMajor`. Use the `geom_bar()` option `position = "fill"` so the bars show relative frequencies. Include an informative title and axis labels using `labs()`. Other colors/formatting are optional.

```{r}

```

## Plot 4: Waffle Plot

Create a waffle plot to show the proportions of `CodedMajor`. Use the `title =` argument in `waffle()` to add an appropriate title.

```{r}

```

## Questions

Your answers should reference specific tables, plots, and/or quantitative values to support your statements. Avoid vague phrases like "the numbers show". Use appropriate statistical vocabulary such as "counts", "frequencies", or "relative frequencies".

1) Overall, what has been the most common major for students in Dr. Thomley's STT3850 sections? Has this been consistently true across all academic years in the dataset?

ANSWER: *erase this and type your answer here*

2) Overall, what has been the most common class year (or years) for students in Dr. Thomley's STT3850 sections? Is this similar across all majors, or do some take the course earlier/later?

ANSWER: *erase this and type your answer here*

3) Using your own knowledge and what is available in the data, can you speculate on a cause or explanation for your findings in Questions 1 and 2? What other information would you like to have to help explain what you observed?

ANSWER: *erase this and type your answer here*

## Reference Materials

* [EDA ⯈ Categorical Data: Tabulating Data](https://stat-jet-asu.github.io/Slides/STT3850/EDACategoricalTables.html#1)
* [EDA ⯈ Categorical Data: Basics of Barplots](https://stat-jet-asu.github.io/Slides/STT3850/EDACategoricalPlots.html#1)
* [EDA ⯈ Categorical Data: Introducing Factors](https://stat-jet-asu.github.io/Slides/STT3850/EDACategoricalFactors.html#1)

***
```{r}
sessionInfo()
```
