---
title: "Measures of Variability"
subtitle: "Companion `R` Code for JB Statistics Video Lesson"
author: "Created by Jill E. Thomley"
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %I:%M %p")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: TRUE
    toc_depth: 4
---

▶ [GO TO VIDEO](https://youtu.be/Cx2tGUze60s)

<hr>

```{r globaloptions, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = NA
)
```

## Packages Used

```{r loadpackages}
library(tidyverse)
library(knitr)
library(kableExtra)
```

## Datasets Used

```{r}
# sample of 4 observations

sample4 <- tibble(data = c(45, 64, 68, 51))
glimpse(sample4)
head(sample4)

# guinea pig survival times

gp <- read_csv("https://raw.githubusercontent.com/STAT-JET-ASU/Datasets/master/JBStatistics/jbstatsguineapigs.csv")
glimpse(gp)
head(gp)
```

## Example with 4 Observations

### Steps in Calculating MAD and SD

Doing the calculations for deviations the long way. Deviations are the distances of each point from the mean. Putting `()` around the code prints out the stored `devs` object. We use `mutate()` to create new columns from the original `data` column.

```{r}
(devs <- sample4 %>%
  mutate(
    deviations = data - mean(data),
    absolute_devs = abs(deviations),
    square_devs = deviations^2
  ))

(sums <- devs %>%
  summarize(
    sum_abs_devs = sum(absolute_devs),
    sum_sqr_devs = sum(square_devs),
    n = n()
  ))

(stats <- sums %>%
  summarize(
    MAD = sum_abs_devs / n,
    s = sqrt(sum_sqr_devs / (n - 1))
  ))
```

### `R` Functions for Summarizing Data

We can do many of the same calculations directly using `R`'s built-in functions.

```{r}
sample4 %>%
  summarize(
    n = n(),
    xbar = mean(data),
    s = sd(data),
    rng = diff(range(data)),
    MAD = sum(abs(data - mean(data))) / n
  )
```

## Example with Guinea Pig Data

Summarize the overall dataset.

```{r}
(gp_stats_all <- gp %>%
  summarize(
    n = n(),
    xbar = mean(survivaldays),
    s = sd(survivaldays),
    rng = diff(range(survivaldays))
  ))
```

We can format the output for presentation using `kable()`.

```{r}
gp_stats_all %>%
  kable(digits = 2) %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    font_size = 14
  )
```

Summarize the data by groups.

```{r}
gp_stats_grp <- gp %>%
  group_by(group) %>%
  summarize(
    n = n(),
    xbar = mean(survivaldays),
    s = sd(survivaldays),
    rng = diff(range(survivaldays))
  )

gp_stats_grp %>%
  kable(digits = 2) %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    font_size = 14
  )
```

## References

Tidyverse: [Create, modify, and delete columns](https://dplyr.tidyverse.org/reference/mutate.html)

Tidyverse: [Summarise each group to fewer rows](https://dplyr.tidyverse.org/reference/summarise.html)

[KableExtra Tutorial](https://rstudio-pubs-static.s3.amazonaws.com/444395_76727eaf9c774fa8bf932bed16500a00.html)

<hr>
```{r}
sessionInfo()
```

